---
title: "p8105_hw2_hm2946.Rmd"
author: "min"
date: "2022-10-04"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
```

## Question 1

```{r}
tran_ent = 
  read_csv(
    "./NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    col_types = cols(Route8 = "c", Route9 = "c",
                     Route10 = "c", Route11 = "c")
  ) %>% 
  janitor::clean_names() %>% 
  select(
     line, station_name, station_latitude, station_longitude, 
     starts_with("Route"), entry, exit_only, vending, 
     entrance_type, ada) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))

tran_ent
```
## First, load the data, and modify the column types;
## Second, clean the data, it contatins line, station_name, station_latitude, station_longitude, Routes, entry, exit_only, vending, entrance_type, ada. The size is 1,868 × 20.
## Last, we convert the entry variable from character to a logical variable, and we know this data is not tidy yet.



## this calculates the number of distinct stations
```{r}
tran_ent %>% 
  select(station_name,line) %>% 
  distinct
```
## 465 distinct stations


## this finds ADA compliant stations
```{r}
tran_ent %>% 
  filter(ada == TRUE) %>% 
  select(station_name,line) %>% 
  distinct
```
## 84 ADA compliant


## this finds proportion of station entrances / exits without vending allow entrance
```{r}
tran_ent %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```
## the proportion is 0.3770492

```{r}
tran_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct

tran_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```
## 60 distnct stations serve the A train, and 17 are ada compliant


## Question 2

```{r}
mr_trash = 
  read_excel("./Trash Wheel Collection Data.xlsx",
             "Mr. Trash Wheel",
             range = "A2:N549") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = as.integer(round(sports_balls))) %>%
  mutate(year = as.double(year)) %>%
  select(-date, weight = weight_tons,
         volume = volume_cubic_yards)

mr_trash
```

```{r}
pro_trash = 
  read_excel("./Trash Wheel Collection Data.xlsx",
             "Professor Trash Wheel",
             range = "A2:M96") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  select(-date, weight = weight_tons,
         volume = volume_cubic_yards)

pro_trash
```

```{r}
mr_trash = 
  mutate(mr_trash, label = "mr") %>% 
  pivot_longer(
    plastic_bottles:sports_balls,
    names_to = "trash_type",
    values_to = "collect_number")
pro_trash = mutate(pro_trash, label = "pro") %>% 
  pivot_longer(
    plastic_bottles:chip_bags,
    names_to = "trash_type",
    values_to = "collect_number")
com_tidy =  bind_rows(mr_trash, pro_trash) %>% 
  janitor::clean_names() %>% 
  select(label, everything()) 

com_tidy
```

## The size of the combined dataset is 4393*9, and it has important variables like year and weight. The total weight of trash collected by Professor Trash Wheel is 1140.72, and the total number of sports balls collected by Mr. Trash Wheel in 2020 is 856.
```{r}
com_tidy %>% 
  filter(label == "pro") %>% 
  select(weight) %>% 
  sum()

com_tidy %>% 
  filter(label == "mr", year == 2020, trash_type == "sports_balls") %>% 
  select(collect_number) %>% 
  sum()
```



## Qeustion 3

```{r}
mon_data = read_csv("./pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(., mon, into = c("year","month","day")) %>% 
  mutate(year = as.integer(year)) %>% 
  mutate(month = as.integer(month)) %>% 
  mutate(day = as.integer(day)) %>% 
  mutate(president = prez_gop + prez_dem) %>% 
  select(-prez_gop, -prez_dem, -day)
mon_data$month = month.abb[mon_data$month]

mon_data
```


```{r}
snp_data = read_csv("./snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(., date, into = c("year","month","day")) %>% 
  mutate(year = as.integer(year)) %>% 
  mutate(month = as.integer(month)) %>%
  select(year, month, close, -day)
snp_data$month = month.abb[snp_data$month]

snp_data
```

```{r}
une_data = read_csv("./unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "unemployment")

une_data
```

```{r}
combined_1 = inner_join(snp_data, mon_data, by = c("year","month"))
combined_2 = inner_join(combined_1, une_data, by = c("year","month"))
```

