---
title: "p8105_hw2_hm2946.Rmd"
author: "min"
date: "2022-10-04"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
```

## Question 1

```{r}
tran_ent = 
  read_csv(
    "./NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    col_types = cols(Route8 = "c", Route9 = "c",
                     Route10 = "c", Route11 = "c")
  ) %>% 
  janitor::clean_names() %>% 
  select(
     line, station_name, station_latitude, station_longitude, 
     starts_with("Route"), entry, exit_only, vending, 
     entrance_type, ada) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))

tran_ent
```
## First, load the data, and modify the column types;
## Second, clean the data, it contatins line, station_name, station_latitude, station_longitude, Routes, entry, exit_only, vending, entrance_type, ada. The size is 1,868 × 20.
## Last, we convert the entry variable from character to a logical variable, and we know this data is not tidy yet.



## this calculates the number of distinct stations
```{r}
tran_ent %>% 
  select(station_name,line) %>% 
  distinct
```
## 465 distinct stations


## this finds ADA compliant stations
```{r}
tran_ent %>% 
  filter(ada == TRUE) %>% 
  select(station_name,line) %>% 
  distinct
```
## 84 ADA compliant


## this finds proportion of station entrances / exits without vending allow entrance
```{r}
tran_ent %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```
## the proportion is 0.3770492

```{r}
tran_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct

tran_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```
## 60 distnct stations serve the A train, and 17 are ada compliant
